### 2015-05-13 whoop.tex
trimming down limitations

Another limitation of \whoop is that it is unable to verify drivers that are designed to be accessed by a single process at a time. This \emph{single-open device}~\cite{corbet2005linux} mode can be enforced by atomically testing (at the beginning of each entry point) a flag that indicates device availability: if the flag is set to true, then the respective entry point executes, else it blocks. Because \whoop performs pair-wise analysis, it over-approximates this flag, and regards a pair of entry points as running concurrently, even if this is not the case. However,~\cite{corbet2005linux} advises against serializing drivers in this way, as it hinders user ingenuity (e.g.\ a user might expect that a device can be accessed concurrently for performance). Because using single-open device mode is considered bad practice, we thus burden the developer with disregarding any related false bug reports.

Statically analyzing system-level software, such as drivers, requires to ``close'' the environment by abstracting away the low level implementation details. In this work, we developed a simple environmental model for the Linux kernel that consists of (nondeterministic) stub functions. A limitation of our model is that it can, and will, ultimately result in decreased precision and, thus, false positives. However, because we currently only focus on finding data races, we can get away with over-approximating a lot of the underlying Linux kernel functionality, without losing too much precision. Enriching our model with domain-specific knowledge to make it more precise is an ongoing manual effort, but requires Linux expertise. We argue that further increasing the precision is orthogonal to the contributions of this paper. Moreover, even if the symbolic lockset analysis results in false positives, \whoop can still use any race-freedom guarantees to significantly speedup a more precise bug-finder, as seen in \S\ref{corral} and \S\ref{evaluation}.

>>>>

Another limitation of \whoop is that it is unable to verify drivers designed to be accessed by only a single process at a time. This \emph{single-open device}~\cite{corbet2005linux} mode can be enforced by atomically testing (at the beginning of an entry point) a flag that indicates device availability: if the flag is set to true, then the checking entry point executes, else it blocks. Because \whoop performs pair-wise analysis, it over-approximates this flag, and thus can falsely report a pair as racy. However,~\cite{corbet2005linux} advises against serializing drivers in this way, as it hinders user ingenuity (e.g.\ a user might expect that a device can be accessed concurrently for performance). Because this serialization is considered bad practice, we burden the developer with disregarding any related false bug reports.

Statically analyzing drivers requires to ``close'' the environment by abstracting away the low-level implementation details. To this end, we developed a simple model for the Linux kernel that consists of (nondeterministic) stub functions. A limitation of our model is that it can, and will, ultimately result in false positives. However, we currently only focus on finding data races, and thus can get away with over-approximating a lot of the underlying kernel functionality, without losing too much precision. Making our model more precise is an ongoing manual effort, but requires Linux expertise. We argue that further work on the model is orthogonal to the contributions of this paper. Also, even if our symbolic analysis results in false positives, \whoop can still use the results to significantly speedup a more precise bug-finder, as seen in \S\ref{corral} and \S\ref{evaluation}.
